Optimizing and Incrementalizing Collection Queries by AST Transformations

### 2019-12-31 aosd13/pepm13/caseStudyImpl.tex
Fix overfull hbox by rephrasing

With these declarations in scope, the Scala compiler rewrites \code{str1 + str2} to
\code{expToStringOps(str1).+}\linebreak\code{(str2)}, which evaluates to \code{StringConcat(str1, str2)} as desired.
The implicit conversion function \code{expToStringOps} is not applicable to \code{Exp[Boolean]} because it explicitly specifies the receiver of the \code{+}-call to have type \code{Exp[String]}.
In other words, expressions like \code{str1 + str2} are now \emph{lifted} on the level of expression trees in a type-safe way. For brevity, we refer to the defined operator as \code{Exp[String].+}.

@@

The compiler will now rewrite \code{str1 + "foo"} to \code{expToStringOps(str1) + pure("foo")}, and
\code{str1 + "foo"}\linebreak\code{+ str2} to \code{expToStringOps(str1) + pure("foo") + str2}.
Different implicit conversions cooperate successfully to lift the expression.
%Since \code{Exp[String].+} requires an argument of type \code{Exp[String]} but \code{"foo"} has type \code{String},
%the compiler wraps \code{"foo"} with a call to \code{pure}. Hence

@@

\smartParagraph{The lifting}
The basic idea is to use the enrich-my-library pattern to lift collection methods from \code{TraversableLike[T, Repr]} to \code{TraversableLikeOps[T, Repr]}:
\begin{lstlisting}
implicit def expToTraversableLikeOps[T, Repr](v: Exp[TraversableLike[T, Repr]]) =
  new TraversableLikeOps[T, Repr](v)
\end{lstlisting}

Subclasses of \code{TraversableLike[T, Repr]} also subclass both \code{Traversable[T]} and \code{Repr}; to take advantage of this during interpretation and optimization, we need to restrict the type of \code{expToTraversableLikeOps}, getting the following conversion:\footnote{Due to type inference bugs, the actual implicit conversion needs to be slightly more complex, to mention \code{T} directly in the argument type. We reported the bug at \url{https://issues.scala-lang.org/browse/SI-5298}.}

@@

\begin{figure}
\begin{center}
\begin{lstlisting}
class TraversableLikeOps[T, Repr <: Traversable[T] with TraversableLike[T, Repr]](
    t: Exp[Repr]) {
  val t: Exp[Repr]

>>>>

With these declarations in scope, the Scala compiler rewrites \code{str1 + str2} to
\code{expToStringOps(}\linebreak\code{str1).+(str2)}, which evaluates to \code{StringConcat(str1, str2)} as desired.
The implicit conversion function \code{expToStringOps} is not applicable to \code{Exp[Boolean]} because it explicitly specifies the receiver of the \code{+}-call to have type \code{Exp[String]}.
In other words, expressions like \code{str1 + str2} are now \emph{lifted} on the level of expression trees in a type-safe way. For brevity, we refer to the defined operator as \code{Exp[String].+}.

@@

The compiler will now rewrite \code{str1 + "foo"} to \code{expToStringOps(str1) + pure("foo")}, and
\code{str1}\linebreak\code{+ "foo" + str2} to \code{expToStringOps(str1) + pure("foo") + str2}.
Different implicit conversions cooperate successfully to lift the expression.
%Since \code{Exp[String].+} requires an argument of type \code{Exp[String]} but \code{"foo"} has type \code{String},
%the compiler wraps \code{"foo"} with a call to \code{pure}. Hence

@@

\smartParagraph{The lifting}
The basic idea is to use the enrich-my-library pattern to lift collection methods from \code{TraversableLike[T, Repr]} to \code{TraversableLikeOps[T, Repr]}:
\begin{lstlisting}
implicit def expToTraversableLikeOps[T, Repr]
    (v: Exp[TraversableLike[T, Repr]]) =
  new TraversableLikeOps[T, Repr](v)
\end{lstlisting}

Subclasses of \code{TraversableLike[T, Repr]} also subclass both \code{Traversable[T]} and \code{Repr}; to take advantage of this subclass relation during interpretation and optimization, we need to restrict the type of \code{expToTraversableLikeOps}, obtaining the following conversion:\footnote{Due to type inference bugs, the actual implicit conversion needs to be slightly more complex, to mention \code{T} directly in the argument type. We reported the bug at \url{https://issues.scala-lang.org/browse/SI-5298}.}

@@

\begin{figure}
\begin{center}
\begin{lstlisting}
class TraversableLikeOps[T,
    Repr <: Traversable[T] with TraversableLike[T, Repr]](t: Exp[Repr]) {
  val t: Exp[Repr]
